Sky Tomography for 3D Aerosol Distribution Recovery

### 2013-07-25 article.tex
bookmark

\usepackage{opex3}
\usepackage{times}
\usepackage{graphicx}
\graphicspath{{./images/}}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}
\usepackage{bm}
\usepackage{enumerate}
\usepackage[]{todonotes} % Add option [disable] to hide all todo notes.

@@

%
% Some new commands I use in this text
%
\newcommand{\OpSphere}{\bm{\mathcal{S}}}
\newcommand{\OpRot}{\bm{\mathcal{R}}}
\newcommand{\OpDistance}{\bm{\mathcal{D}}}
\newcommand{\OpCumsum}{\bm{\mathcal{C}}}
\newcommand{\OpInt}{\bm{\mathcal{I}}}
\newcommand{\OpCamera}{\bm{\mathcal{P}}}
\newcommand{\OpDiag}[1]{\mathrm{diag}\left\{#1\right\}}
\newcommand{\Grad}[1]{\bm{\triangledown_{#1}}}
\newcommand{\argmin}{\mathrm{arg}\min}
\newcommand{\curly}[1]{\left\{#1\right\}}
\newcommand{\roundy}[1]{\left(#1\right)}
\newcommand{\recty}[1]{\left[#1\right]}
\newcommand{\PartDeriv}[2]{\frac{\partial{#1}}{\partial{#2}}}
\newcommand{\vect}[1]{\bm{#1}}
\newcommand{\mat}[1]{\bm{#1}}
\newcommand{\transpose}[1]{{#1}^\intercal}
\newcommand{\derivsym}[1]{\,d{#1}}
\newcommand{\yoavcomment}[1]{}
%\renewcommand{\yoavcomment}[1]{#1} % Comment to remove images

@@

\begin{abstract}
  Aerosols effect climate, health and aviation.  Currently, their
  retrieval assumes a plane-parallel atmosphere and solely vertical
  radiative transfer. We propose a principle to recover the atmosphere
  as it really is: a three dimensional (3D), volumetric, transparent
  huge object. Tomography is the key to recovery. The process involves
  acquiring the sky lightfield by multi-view wide-angle photography on
  a very large scale. The tomography model is distinct, as the light
  source is unidirectional and uncontrolled, while off-axis scattering
  dominates the images.  We formulate an image formation model based
  on 3D radiative transfer. Model inversion is done using optimization
  methods, exploiting a closed-form gradient which we derive for the
  model-fit cost function. The problem is complex but tractable and
  parallelizable, eventually estimating the 3D aerosol distribution.
\end{abstract}

@@

We do not deal with clouds here. Clouds scatter sunlight very strongly
and are easily observed. Hence they are directly recovered using
common stereoscopic tools, similarly to opaque surfaces~\cite{seiz}.
On the other hand, aerosols have much more subtle visual effects and
are thus more difficult to recover. We focus on them.  In remote
sensing~\cite{diner}, imaging through air is associated with {\em
  atmospheric correction}, based on {\em aerosol retrieval}. In this
discipline, the atmosphere is often assumed to be plane-parallel, for
simplicity, using only 1D vertical radiative transfer. Consequently
state-of-the-art aerosol retrieval is done in distinct large lateral
blocks~\cite{Martonchikc} with limited height
resolution~\cite{kalashnikova}. We, however, seek high resolution 3D
recovery. Computational photography offers new ways to peer into
media. In addition to active descattering~\cite{fuchs,kim,Ng1948} and
recovery of refractive transparent objects~\cite{ihrke}, overcoming
the atmosphere~\cite{Joshi2010,Zhu2013} to view ground objects was
studied in horizontal, fixed
views~\cite{fattal,he,Chen,kratz,narasimhan2,oakley,Hschechner2,tan}. In
this paper, however, the medium itself, at all relevant altitudes is
the object of interest.

@@

\noindent {\bf Extinction}: Sun rays (SR) irradiate a small volume
that includes particles of a certain type.  Each particle has an {\em
  extinction cross section} $\sigma$ (units \si{\meter\squared}) for
interacting with the irradiance. The volume includes many particles of
this type. The denser the volume, the stronger is the extinction.  The
density of the particles is $n$ (units
\si[sticky-per]{\per\cubic\metre}). Per unit volume, the {\em
  extinction coefficient} $\beta$ (units \si{\per\meter}) is
\begin{align}
  \beta= \sigma n \;.
  \label{eq:extinctc}
\end{align}
The volume has infinitesimal length $dl$. Then, the relative portion
of extinct SR irradiance is the unitless differential {\em optical
  depth}
\begin{align}
  d\tau= \beta dl=\sigma n dl \;.
  \label{eq:extinct}
\end{align}
The optical depth aggregates in extended propagation:
\begin{align}
  \tau= \int d\tau=\int \beta dl=\int \sigma n dl \;.
  \label{eq:tau}
\end{align}
Through an attenuating volume, the {\em transmittance} exponentially
decays with the optical depth:
\begin{align}
  t=\exp^{-\tau}
  \label{eq:beer-lambert}
\end{align}

\noindent {\bf Scattering}: Interaction of a single particle with the
irradiance is by absorption and scattering. The weight of scattering
(to all directions), relative to the total extinction is given by the
unitless {\em single scattering albedo} $\varpi$ of the particle. The
{\em scattering coefficient} $\alpha$ (units \si{\per\meter}) is
\begin{align}
  \alpha= \varpi\beta=\varpi \sigma n \;.
  \label{eq:alph}
\end{align}
Scattering is generally anisotropic. Its angular distribution is
determined by a {\em phase function} $P$. Part of the light scatters
towards a camera's line of sight (LOS), as illustrated in
\cref{fig:groundgrid}. The angle between the SR and LOS is the
scattering angle $\Phi^{\rm scatter}$. The phase function $P(\Phi^{\rm
  scatter})$ is normalized: its integral over all solid angles is
unit. The {\em angular scattering coefficient} $\tilde\alpha$ (units
\si[sticky-per]{\per\meter\steradian}) is
\begin{align}
  \tilde\alpha(\Phi^{\rm scatter}) = \varpi\beta P(\Phi^{\rm scatter})
  = \varpi \sigma n P(\Phi^{\rm scatter}).
  \label{eq:alphabasic}
\end{align}

The phase function is often approximated by a parametric expression,
as the Henyey-Greenstein~\cite{Cornette1995} function,
\begin{align}
  P_g(\Phi^{\rm scatter})\approx \frac{3} {8\pi} \frac{(1 -
    g^2)(1+\mu^2)} {(2+g^2)(1 + g^2 - 2g\mu)^\frac{3}{2}}
  \label{eq:aerosol_scatter}
\end{align}
where $g$ is an {\em anisotropy parameter} and
\begin{align}
  \mu\equiv \cos \Phi^{\rm scatter}.
  \label{eq:mu}
\end{align}

\noindent {\bf Air molecules}: Scattering by air molecules follows the
{\em Rayleigh} model:
\begin{align}
  P^{\rm air}\left[\mu(k)\right] &\approx \frac{3}{16\pi}(1+\mu^2)
  \label{eq:rayleighP}
\end{align}
and $\varpi^{\rm air}$=1. Air molecular density $n^{\rm air}$ falls
off approximately exponentially with altitude, with a
characteristic~\cite{Levi1980} falloff height $H^\mathrm{air}=8\
\si{\km}$. Consequently~\cite{Levi1980}, the coefficients for
extinction and scattering by air molecules are modeled by
\begin{align}
  \alpha^{\rm air}(h, \lambda)=\beta^{\rm air}(h, \lambda) \approxeq
  \frac{1.09 \times 10^{-3}}{\lambda^4}
  \exp(-h/H^\mathrm{air}) %\left[-\frac{h}{H^\mathrm{air}}\right].
  \label{eq:rayleighbeta}
\end{align}

@@

We believe that high resolution atmospheric recovery can be achieved
by a large set of ground-based wide-angle~\cite{Cossairt2011,
  Peleg01omnistereo:panoramic, Koppal:2011:WMS:2191740.2192152}
sky-cameras, looking upward and capturing the sky lightfield.  This
paper takes the first step, of formulating this principle and
demonstrating feasibility using realistic simulations. The formulation
of this paper can also apply to multiview images that are captured
from satellites, for future spaceborne instruments of this capacity.

Our reconstruction algorithm is based on the
Single-Scattering model (\cref{sec:single-scatt-model}). The
Single-Scattering approximation enables stating an analytic solution
to the Radiative Transfer in a closed form. This closed form is
beneficial to us as it enables us to invert the model and thus recover
the underlying atmosphere (\cref{sec:inverse-problem}). We validate
our reconstruction method by applying it on 'Realistic' images.
We simulate image acquisition using Monte Carlo (MC) simulations
(\cref{sec:monte-carlo-simul}).  The monte carlo method models
multiple scattering and therefore enables simulating realistic
scenarios.

@@

The recovery task is phrased as optimization of an objective function,
that fits the image-formation model (\cref{eq:bigIc,eq:bigIA}) to
measured photographs $\{ {\bm i}^{\rm measured}_c\}_{c=1}^{N_{\rm
    views}}$:
\begin{equation}
  \label{eq:minobjectiveA}
  \hat{\bm n} =
  \argmin_{{\bm n}} F({\bm n})
\end{equation}
where
\begin{equation}
  \label{eq:objectiveA}
  F({\bm n})
  = \sum_{c=1}^{N_{\rm views}}
  \| {\bm i}^{\rm measured}_c - {\bm i}_c({\bm n})\|^2_2 + \eta \cdot \Psi({\bm n})
\end{equation}
Here $\Psi({\bm n})$ is a regularization term that expresses some
prior knowledge on the solution and $\eta$ is the regularization
coefficient.  We enforce smoothness of the solution by using the
\emph{Laplacian} operator.
\begin{equation}
  \label{eq:regularizer}
  \Psi({\bm n}) = \| \nabla^2{\bm n}\|^2_2
\end{equation}
We solve~\cref{eq:minobjectiveA} using standard optimization tools.
The gradient of $F$ with respect to ${\bm n}$ is
\begin{align}
  \Grad{{\bm n}} F = -2\sum_{c=1}^{N_{\rm views}}
  \transpose{\left[{\bm J}_{{\bm i}_c}({\bm n})\right]} [{\bm i}^{\rm
    measured}_c - {\bm i}_c({\bm n})] + 2 \cdot \eta \cdot
  \nabla^2\nabla^2{\bm n} \;.
  \label{eq:gradient1}
\end{align}
Here the matrix ${\bm J}_{{\bm i}_c}({\bm n})$ is the Jacobian of the
vector ${\bm i}_c$ with respect to ${\bm n}$, and we make use of the
symmetry of the \emph{Laplacian} operator. A close-form expression of
our Jacobian is
\begin{align}
  {\bm J}_{{\bm i}_c}({\bm n}) &= L^{\rm TOA}({\bm A}-{\bm B}){\bm C}
  ,
  \label{eq:gradient2}
\end{align}
where
\begin{align}
  \label{eq:gradient3}
  {\bm A} &= \varpi^{\rm aerosol}\sigma^{\rm aerosol}
  \OpDiag{ P_g^{\rm aerosol}(\vect{\Phi}^{\rm scatter}_c)} \nonumber\\
  {\bm B} &= {\sigma}^{\rm aerosol}\transpose{{\bm D}_c} \nonumber\\
  & \;\;\; \OpDiag{[\tilde{\vect{\alpha}}^{\rm air}_c + \varpi^{\rm
      aerosol}\sigma^{\rm aerosol} P^{\rm aerosol}_g(\vect{\Phi}^{\rm
      scatter}_c) \odot{\bm n}
    ]} \nonumber \\
  {\bm C} &= \OpDiag{\exp[-(\vect{\tau}_c^{\rm air} + {\sigma}^{\rm
      aerosol} {\bm D}_c {\bm n})]} \transpose{{\vect{\Pi}}_c}
\end{align}
For a detailed derivation of \cref{eq:gradient2,eq:gradient3}, we
refer the reader to \cref{sec:gradient-derivation}.

@@

Monte Carlo (MC) methods enable rigorous modeling of photon
propagation in the atmosphere. The radiative transfer is expressed as
probability distributions which describe the distance travelled by the
photon between scatterings and the angle of deflection at each
scatter.  This enables simulating multiple scattering and produce
realistic results.  MC simulations can be made arbitrarily accurate by
increasing the number of photons.  Another advantage of the MC methods
is that they are highly parallelable.

>>>>

\usepackage{opex3}
\usepackage{times}
\usepackage{graphicx}
\graphicspath{{./images/}}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}
\usepackage{bm}
\usepackage{enumerate}
\usepackage[]{todonotes} % Add option [disable] to hide all todo notes.
\usepackage{xfrac}

@@

%
% Some new commands I use in this text
%
\newcommand{\OpSphere}{\bm{\mathcal{S}}}
\newcommand{\OpRot}{\bm{\mathcal{R}}}
\newcommand{\OpDistance}{\bm{\mathcal{D}}}
\newcommand{\OpCumsum}{\bm{\mathcal{C}}}
\newcommand{\OpInt}{\bm{\mathcal{I}}}
\newcommand{\OpCamera}{\bm{\mathcal{P}}}
\newcommand{\Mask}{\bm{M}}
\newcommand{\OpDiag}[1]{\mathrm{diag}\left\{#1\right\}}
\newcommand{\Grad}[1]{\bm{\triangledown_{#1}}}
\newcommand{\argmin}{\mathrm{arg}\min}
\newcommand{\curly}[1]{\left\{#1\right\}}
\newcommand{\roundy}[1]{\left(#1\right)}
\newcommand{\recty}[1]{\left[#1\right]}
\newcommand{\PartDeriv}[2]{\frac{\partial{#1}}{\partial{#2}}}
\newcommand{\vect}[1]{\bm{#1}}
\newcommand{\mat}[1]{\bm{#1}}
\newcommand{\transpose}[1]{{#1}^\intercal}
\newcommand{\derivsym}[1]{\,d{#1}}
\newcommand{\yoavcomment}[1]{}
\renewcommand{\yoavcomment}[1]{#1} % Comment to remove images

@@

\begin{abstract}
  Aerosols effect climate, health and aviation.  Currently, their
  retrieval assumes a plane-parallel atmosphere and solely vertical
  radiative transfer (RT). We propose a principle to recover the atmosphere
  as it really is: a three dimensional (3D), volumetric, transparent
  huge object. Tomography is the key to recovery. The process involves
  acquiring the sky lightfield by multi-view wide-angle photography on
  a very large scale. The tomography model is distinct, as the light
  source is unidirectional and uncontrolled, while off-axis scattering
  dominates the images.  We formulate an image formation model based
  on 3D RT. Model inversion is done using optimization
  methods, exploiting a closed-form gradient which we derive for the
  model-fit cost function. The problem is complex but tractable and
  parallelizable, eventually estimating the 3D aerosol distribution.
\end{abstract}

@@

We do not deal with clouds here. Clouds scatter sunlight very strongly
and are easily observed. Hence they are directly recovered using
common stereoscopic tools, similarly to opaque surfaces~\cite{seiz}.
On the other hand, aerosols have much more subtle visual effects and
are thus more difficult to recover. We focus on them.  In remote
sensing~\cite{diner}, imaging through air is associated with {\em
  atmospheric correction}, based on {\em aerosol retrieval}. In this
discipline, the atmosphere is often assumed to be plane-parallel, for
simplicity, using only 1D vertical RT. Consequently
state-of-the-art aerosol retrieval is done in distinct large lateral
blocks~\cite{Martonchikc} with limited height
resolution~\cite{kalashnikova}. We, however, seek high resolution 3D
recovery. Computational photography offers new ways to peer into
media. In addition to active descattering~\cite{fuchs,kim,Ng1948} and
recovery of refractive transparent objects~\cite{ihrke}, overcoming
the atmosphere~\cite{Joshi2010,Zhu2013} to view ground objects was
studied in horizontal, fixed
views~\cite{fattal,he,Chen,kratz,narasimhan2,oakley,Hschechner2,tan}. In
this paper, however, the medium itself, at all relevant altitudes is
the object of interest.

@@

\noindent {\bf Extinction}: Sun rays (SR) irradiate a small volume
that includes particles of a certain type.  Each particle has an {\em
  extinction cross section} $\sigma$ (units \si{\meter\squared}) for
interacting with the irradiance. The density of the particles is $n$
(units \si[sticky-per]{\per\cubic\metre}). Per unit volume, the {\em
  extinction coefficient} $\beta$ is
\begin{align}
  \beta= \sigma n \;.
  \label{eq:extinctc}
\end{align}
The volume has infinitesimal length $dl$. Then, the relative portion
of extinct SR irradiance is the unitless differential {\em optical
  depth}
\begin{align}
  d\tau= \beta dl=\sigma n dl \;.
  \label{eq:extinct}
\end{align}
The optical depth aggregates in extended propagation:
\begin{align}
  \tau= \int d\tau=\int \beta dl=\int \sigma n dl \;.
  \label{eq:tau}
\end{align}
Through an attenuating volume, the {\em transmittance} exponentially
decays with the optical depth:
\begin{align}
  t=\exp(-\tau)
  \label{eq:beer-lambert}
\end{align}

\noindent {\bf Scattering}: Interaction of a single particle with the
irradiance is by absorption and scattering. The weight of scattering
(to all directions), relative to the total extinction is given by the
unitless {\em single scattering albedo} $\varpi$ of the particle. The
{\em scattering coefficient} $\alpha$ is
\begin{align}
  \alpha= \varpi\beta=\varpi \sigma n \;.
  \label{eq:alph}
\end{align}
The angular distribution of scattering is determined by a {\em phase
  function} $P$. Part of the light scatters towards a camera's LOS, as
illustrated in \cref{fig:groundgrid}. The angle between the SR and LOS
is the scattering angle $\Phi^{\rm scatter}$. The phase function
$P(\Phi^{\rm scatter})$ is normalized: its integral over all solid
angles is unit. The {\em angular scattering coefficient}
$\tilde\alpha$ (units \si[sticky-per]{\per\meter\steradian}) is
\begin{align}
  \tilde\alpha(\Phi^{\rm scatter}) = \varpi\beta P(\Phi^{\rm scatter})
  = \varpi \sigma n P(\Phi^{\rm scatter}).
  \label{eq:alphabasic}
\end{align}

The phase function is often approximated by a parametric expression,
as the Henyey-Greenstein~\cite{Cornette1995} function,
\begin{align}
  P_g(\Phi^{\rm scatter})\approx \sfrac{3}{8\pi} (1 - g^2)(1+\mu^2)
  (2+g^2)^{-1}(1 + g^2 - 2g\mu)^{\sfrac{-3}{2}}
  \label{eq:aerosol_scatter}
\end{align}
where $g$ is an {\em anisotropy parameter} and $\mu\equiv \cos \Phi^{\rm scatter}$.
\\

\noindent {\bf Air molecules}: Scattering by air molecules follows the
{\em Rayleigh} model:
\begin{align}
  P^{\rm air}\left[\mu(k)\right] &\approx \sfrac{3}{16\pi}(1+\mu^2)
  \label{eq:rayleighP}
\end{align}
and $\varpi^{\rm air}$=1. Air molecular density $n^{\rm air}$ falls
off approximately exponentially with altitude, with a
characteristic~\cite{Levi1980} falloff height $H^\mathrm{air}=8\
\si{\km}$. Consequently~\cite{Levi1980}, the coefficients for
extinction and scattering by air molecules are modeled by
\begin{align}
  \alpha^{\rm air}(h, \lambda)=\beta^{\rm air}(h, \lambda) \approxeq
  1.09 \times 10^{-3}\lambda^{-4}
  \exp(-h/H^\mathrm{air}) %\left[-\frac{h}{H^\mathrm{air}}\right].
  \label{eq:rayleighbeta}
\end{align}

@@

We believe that high resolution atmospheric recovery can be achieved
by a set of ground-based wide-angle~\cite{Cossairt2011,
  Peleg01omnistereo:panoramic, Koppal:2011:WMS:2191740.2192152}
sky-cameras, looking upward, performing integral imaging of the sky.
This paper takes the first step, of formulating this principle and
demonstrating feasibility using realistic simulations. The formulation
of this paper can also apply to multiview images that are captured
from satellites, for future spaceborne instruments of this capacity.

We simulate RT through the atmosphere using two models. The first is
based on Monte Carlo (MC) methods (\cref{sec:monte-carlo-simul}) and
models Multiple-Scattering. Therefore it enables simulating realistic
scenarios.  The second model uses the Single-Scattering approximation
(\cref{sec:single-scatt-model}). It is less accurate than the MC model
but it enables stating an analytic solution to the RT in a closed
form. This closed form is beneficial to us as it enables us to invert
the model and thus recover the underlying atmosphere
(\cref{sec:inverse-problem}). To mimic realistic scenarios, we apply
the reconstruction algorithm on images simulated by the MC model.

@@

The recovery task is phrased as optimization of an objective function,
that fits the image-formation model (\cref{eq:bigIc,eq:bigIA}) to
measured photographs $\{ {\bm i}^{\rm measured}_c\}_{c=1}^{N_{\rm
    views}}$:
\begin{equation}
  \label{eq:minobjectiveA}
  \hat{\bm n} =
  \argmin_{{\bm n}} F({\bm n})
\end{equation}
where
\begin{equation}
  \label{eq:objectiveA}
  F({\bm n})
  = \sum_{c=1}^{N_{\rm views}}
  \| \Mask \roundy{{\bm i}^{\rm measured}_c - {\bm i}_c({\bm n})}\|^2_2 + \eta \cdot \Psi({\bm n})
\end{equation}
Here $\Psi({\bm n})$ is a regularization term that expresses some
prior knowledge on the solution and $\eta$ is the regularization
coefficient.  We enforce smoothness of the solution by using the
\emph{Laplacian} operator.
\begin{equation}
  \label{eq:regularizer}
  \Psi({\bm n}) = \| \nabla^2{\bm n}\|^2_2
\end{equation}
We use $\Mask$ as a weight map to mask the influence of the sun and
the horizon. These regions are prone to dominant multiple scattering.
We solve~\cref{eq:minobjectiveA} using standard optimization tools.
The gradient of $F$ with respect to ${\bm n}$ is
\begin{align}
  \Grad{{\bm n}} F = -2\sum_{c=1}^{N_{\rm views}}
  \transpose{\left[{\bm J}_{{\bm i}_c}({\bm n})\right]} [{\bm i}^{\rm
    measured}_c - {\bm i}_c({\bm n})] + 2 \cdot \eta \cdot
  \nabla^2\nabla^2{\bm n} \;.
  \label{eq:gradient1}
\end{align}
Here the matrix ${\bm J}_{{\bm i}_c}({\bm n})$ is the Jacobian of the
vector ${\bm i}_c$ with respect to ${\bm n}$, and we make use of the
symmetry of the \emph{Laplacian} operator. A close-form expression of
our Jacobian is
\begin{align}
  {\bm J}_{{\bm i}_c}({\bm n}) &= L^{\rm TOA}({\bm A}-{\bm B}){\bm C}
  ,
  \label{eq:gradient2}
\end{align}
where
\begin{align}
  \label{eq:gradient3}
  {\bm A} &= \varpi^{\rm aerosol}\sigma^{\rm aerosol}
  \OpDiag{ P_g^{\rm aerosol}(\vect{\Phi}^{\rm scatter}_c)} \nonumber\\
  {\bm B} &= {\sigma}^{\rm aerosol}\transpose{{\bm D}_c} \nonumber\\
  & \;\;\; \OpDiag{[\tilde{\vect{\alpha}}^{\rm air}_c + \varpi^{\rm
      aerosol}\sigma^{\rm aerosol} P^{\rm aerosol}_g(\vect{\Phi}^{\rm
      scatter}_c) \odot{\bm n}
    ]} \nonumber \\
  {\bm C} &= \OpDiag{\exp[-(\vect{\tau}_c^{\rm air} + {\sigma}^{\rm
      aerosol} {\bm D}_c {\bm n})]} \transpose{{\vect{\Pi}}_c}
\end{align}
For a detailed derivation of \cref{eq:gradient2,eq:gradient3}, we
refer the reader to \cref{sec:gradient-derivation}.

@@

Monte Carlo (MC) methods enable rigorous modeling of photon
propagation in the atmosphere. The radiative transfer is expressed as
probability distributions which describe the distance travelled by the
photon between scatterings and the angle of deflection at each
scatter.  This enables simulating multiple scattering and produce
realistic results.  MC simulations can be made arbitrarily accurate by
increasing the number of photons.  Another advantage of the MC methods
is that they are highly parallelizable.
